\chapter{Global parameters}
Parameters in the "global" section are process-wide and often OS-specific. They
are generally set once for all and do not need being changed once correct. Some
of them have command-line equivalents.

The following keywords are supported in the "global" section:

\begin{itemize}
\item Process management and security
  \begin{itemize}
    \item[-] chroot
    \item[-] daemon
    \item[-] gid
    \item[-] group
    \item[-] log
    \item[-] log-send-hostname
    \item[-] nbproc
    \item[-] pidfile
    \item[-] uid
    \item[-] ulimit-n
    \item[-] user
    \item[-] stats
    \item[-] node
    \item[-] description
    \item[-] unix-bind
  \end{itemize}
\item Performance tuning
  \begin{itemize}
    \item[-] maxconn
    \item[-] maxconnrate
    \item[-] maxpipes
    \item[-] noepoll
    \item[-] nokqueue
    \item[-] nopoll
    \item[-] nosepoll
    \item[-] nosplice
    \item[-] spread-checks
    \item[-] tune.bufsize
    \item[-] tune.chksize
    \item[-] tune.http.maxhdr
    \item[-] tune.maxaccept
    \item[-] tune.maxpollevents
    \item[-] tune.maxrewrite
    \item[-] tune.pipesize
    \item[-] tune.rcvbuf.client
    \item[-] tune.rcvbuf.server
    \item[-] tune.sndbuf.client
    \item[-] tune.sndbuf.server
  \end{itemize}
\item Debugging
  \begin{itemize}
    \item[-] debug
    \item[-] quiet
  \end{itemize}
\end{itemize}

\section{Process management and security}

\subsubsection[chroot]{chroot <jail dir>}
  Changes current directory to <jail dir> and performs a chroot() there before
  dropping privileges. This increases the security level in case an unknown
  vulnerability would be exploited, since it would make it very hard for the
  attacker to exploit the system. This only works when the process is started
  with superuser privileges. It is important to ensure that <jail\_dir> is both
  empty and unwritable to anyone.
  
\subsubsection[daemon]{daemon}
  Makes the process fork into background. This is the recommended mode of
  operation. It is equivalent to the command line "-D" argument. It can be
  disabled by the command line "-db" argument.
  
\subsubsection[gid]{gid <number>}
  Changes the process' group ID to <number>. It is recommended that the group
  ID is dedicated to HAProxy or to a small set of similar daemons. HAProxy must
  be started with a user belonging to this group, or with superuser privileges.
  See also "group" and "uid".
  
\subsubsection[group]{group <group name>}
  Similar to "gid" but uses the GID of group name <group name> from /etc/group.
  See also "gid" and "user".

\subsubsection[log]{log <address> <facility> [max level [min level]]}
  Adds a global syslog server. Up to two global servers can be defined. They
  will receive logs for startups and exits, as well as all logs from proxies
  configured with "log global".

  <address> can be one of:
  
  \begin{itemize}
    \item[-] An IPv4 address optionally followed by a colon and a UDP port. If
          no port is specified, 514 is used by default (the standard syslog
          port).

    \item[-] An IPv6 address followed by a colon and optionally a UDP port. If
          no port is specified, 514 is used by default (the standard syslog
          port).

    \item[-] A filesystem path to a UNIX domain socket, keeping in mind
          considerations for chroot (be sure the path is accessible inside
          the chroot) and uid/gid (be sure the path is appropriately
          writeable).
  \end{itemize}
  
  <facility> must be one of the 24 standard syslog facilities:
  
  \vspace{5mm}
  \textbf{
  \begin{tabular}{llllllll}
    kern   & user   & mail   & daemon & auth   & syslog & lpr    & news \\
    uucp   & cron   & auth2  & ftp    & ntp    & audit  & alert  & cron2 \\
    local0 & local1 & local2 & local3 & local4 & local5 & local6 & local7
  \end{tabular}
  }
  \vspace{5mm}
    
  An optional level can be specified to filter outgoing messages. By default,
  all messages are sent. If a maximum level is specified, only messages with a
  severity at least as important as this level will be sent. An optional minimum
  level can be specified. If it is set, logs emitted with a more severe level
  than this one will be capped to this level. This is used to avoid sending
  "emerg" messages on all terminals on some default syslog configurations.
  
  Eight levels are known:
  
  \vspace{3mm}
  \textbf {
  \begin{tabular}{llllllll}
          emerg & alert & crit & err & warning & notice & info & debug
  \end{tabular}
  }
  
\subsubsection[log-send-hostname]{log-send-hostname [<string>]}
  Sets the hostname field in the syslog header. If optional "string" parameter
  is set the header is set to the string contents, otherwise uses the hostname
  of the system. Generally used if one is not relaying logs through an
  intermediate syslog server or for simply customizing the hostname printed in
  the logs.

\subsubsection[log-tag]{log-tag <string>}
  Sets the tag field in the syslog header to this string. It defaults to the
  program name as launched from the command line, which usually is "haproxy".
  Sometimes it can be useful to differentiate between multiple processes
  running on the same host.

\subsubsection[nbproc]{nbproc <number>}
  Creates <number> processes when going daemon. This requires the "daemon"
  mode. By default, only one process is created, which is the recommended mode
  of operation. For systems limited to small sets of file descriptors per
  process, it may be needed to fork multiple daemons. \emph{Using multiple processes
  is harder to debug and is really discouraged.}
  
  See also "daemon".

\subsubsection[pidfile]{pidfile <pidfile>}
  Writes pids of all daemons into file <pidfile>. This option is equivalent to
  the "-p" command line argument. The file must be accessible to the user
  starting the process. See also "daemon".

\subsubsection[stats socket]{stats socket <path> [{uid | user} <uid>] [{gid | group} <gid>] [mode <mode>] [level <level>]}

  Creates a UNIX socket in stream mode at location <path>. Any previously
  existing socket will be backed up then replaced. Connections to this socket
  will return various statistics outputs and even allow some commands to be
  issued. Please consult section 9.2 "Unix Socket commands" for more details.
  
  An optional "level" parameter can be specified to restrict the nature of
  the commands that can be issued on the socket:
  
  \begin{itemize}
    \item[-] "user" is the least privileged level; only non-sensitive stats can be
      read, and no change is allowed. It would make sense on systems where it
      is not easy to restrict access to the socket.

    \item[-] "operator" is the default level and fits most common uses. All data can
      be read, and only non-sensitive changes are permitted (eg: clear max
      counters).

    \item[-] "admin" should be used with care, as everything is permitted (eg: clear
      all counters).
  \end{itemize}
  
  On platforms which support it, it is possible to restrict access to this
  socket by specifying numerical IDs after "uid" and "gid", or valid user and
  group names after the "user" and "group" keywords. It is also possible to
  restrict permissions on the socket by passing an octal value after the "mode"
  keyword (same syntax as chmod). Depending on the platform, the permissions on
  the socket will be inherited from the directory which hosts it, or from the
  user the process is started with.

\subsubsection[stats timeout]{stats timeout <timeout, in milliseconds>}
  The default timeout on the stats socket is set to 10 seconds. It is possible
  to change this value with "stats timeout". The value must be passed in
  milliseconds, or be suffixed by a time unit among { us, ms, s, m, h, d }.

\subsubsection[stats maxconn]{stats maxconn <connections>}
  By default, the stats socket is limited to 10 concurrent connections. It is
  possible to change this value with "stats maxconn".

\subsubsection[uid]{uid <number>}
  Changes the process' user ID to <number>. It is recommended that the user ID
  is dedicated to HAProxy or to a small set of similar daemons. HAProxy must
  be started with superuser privileges in order to be able to switch to another
  one. See also "gid" and "user".

\subsubsection[ulimit-n]{ulimit-n <number>}
  Sets the maximum number of per-process file-descriptors to <number>. By
  default, it is automatically computed, so it is recommended not to use this
  option.

\subsubsection[unix-bind]{unix-bind [ prefix <prefix> ] [ mode <mode> ] [ user <user> ] [ uid <uid> ] [ group <group> ] [ gid <gid> ]}

  Fixes common settings to UNIX listening sockets declared in "bind" statements.
  This is mainly used to simplify declaration of those UNIX sockets and reduce
  the risk of errors, since those settings are most commonly required but are
  also process-specific. The <prefix> setting can be used to force all socket
  path to be relative to that directory. This might be needed to access another
  component's chroot. Note that those paths are resolved before haproxy chroots
  itself, so they are absolute. The <mode>, <user>, <uid>, <group> and <gid>
  all have the same meaning as their homonyms used by the "bind" statement. If
  both are specified, the "bind" statement has priority, meaning that the
  "unix-bind" settings may be seen as process-wide default settings.

\subsubsection[user]{user <user name>}
  Similar to "uid" but uses the UID of user name <user name> from /etc/passwd.
  See also "uid" and "group".

\subsubsection[node]{node <name>}
  Only letters, digits, hyphen and underscore are allowed, like in DNS names.

  This statement is useful in HA configurations where two or more processes or
  servers share the same IP address. By setting a different node-name on all
  nodes, it becomes easy to immediately spot what server is handling the
  traffic.

\subsubsection[description]{description <text>}
  Add a text that describes the instance.

  Please note that it is required to escape certain characters (\# for example)
  and this text is inserted into a html page so you should avoid using
  "<" and ">" characters.

\section{Performance tuning}

\subsubsection[maxconn]{maxconn <number>}
  Sets the maximum per-process number of concurrent connections to <number>. It
  is equivalent to the command-line argument "-n". Proxies will stop accepting
  connections when this limit is reached. The "ulimit-n" parameter is
  automatically adjusted according to this value. See also "ulimit-n".

\subsubsection[maxconnrate]{maxconnrate <number>}
  Sets the maximum per-process number of connections per second to <number>.
  Proxies will stop accepting connections when this limit is reached. It can be
  used to limit the global capacity regardless of each frontend capacity. It is
  important to note that this can only be used as a service protection measure,
  as there will not necessarily be a fair share between frontends when the
  limit is reached, so it's a good idea to also limit each frontend to some
  value close to its expected share. Also, lowering tune.maxaccept can improve
  fairness.

\subsubsection[maxpipes]{maxpipes <number>}
  Sets the maximum per-process number of pipes to <number>. Currently, pipes
  are only used by kernel-based tcp splicing. Since a pipe contains two file
  descriptors, the "ulimit-n" value will be increased accordingly. The default
  value is maxconn/4, which seems to be more than enough for most heavy usages.
  The splice code dynamically allocates and releases pipes, and can fall back
  to standard copy, so setting this value too low may only impact performance.

\subsubsection[noepoll]{noepoll}
  Disables the use of the "epoll" event polling system on Linux. It is
  equivalent to the command-line argument "-de". The next polling system
  used will generally be "poll". See also "nosepoll", and "nopoll".

\subsubsection[nokqueue]{nokqueue}
  Disables the use of the "kqueue" event polling system on BSD. It is
  equivalent to the command-line argument "-dk". The next polling system
  used will generally be "poll". See also "nopoll".

\subsubsection[nopoll]{nopoll}
  Disables the use of the "poll" event polling system. It is equivalent to the
  command-line argument "-dp". The next polling system used will be "select".
  It should never be needed to disable "poll" since it's available on all
  platforms supported by HAProxy. See also "nosepoll", and "nopoll" and
  "nokqueue".

\subsubsection[nosepoll]{nosepoll}
  Disables the use of the "speculative epoll" event polling system on Linux. It
  is equivalent to the command-line argument "-ds". The next polling system
  used will generally be "epoll". See also "nosepoll", and "nopoll".

\subsubsection[nosplice]{nosplice}
  Disables the use of kernel tcp splicing between sockets on Linux. It is
  equivalent to the command line argument "-dS".  Data will then be copied
  using conventional and more portable recv/send calls. Kernel tcp splicing is
  limited to some very recent instances of kernel 2.6. Most versions between
  2.6.25 and 2.6.28 are buggy and will forward corrupted data, so they must not
  be used. This option makes it easier to globally disable kernel splicing in
  case of doubt. See also "option splice-auto", "option splice-request" and
  "option splice-response".

\subsubsection[spread-checks]{spread-checks <0..50, in percent>}
  Sometimes it is desirable to avoid sending health checks to servers at exact
  intervals, for instance when many logical servers are located on the same
  physical server. With the help of this parameter, it becomes possible to add
  some randomness in the check interval between 0 and +/- 50%. A value between
  2 and 5 seems to show good results. The default value remains at 0.

\subsubsection[tune.bufsize]{tune.bufsize <number>}
  Sets the buffer size to this size (in bytes). Lower values allow more
  sessions to coexist in the same amount of RAM, and higher values allow some
  applications with very large cookies to work. The default value is 16384 and
  can be changed at build time. It is strongly recommended not to change this
  from the default value, as very low values will break some services such as
  statistics, and values larger than default size will increase memory usage,
  possibly causing the system to run out of memory. At least the global maxconn
  parameter should be decreased by the same factor as this one is increased.

\subsubsection[tune.chksize]{tune.chksize <number>}
  Sets the check buffer size to this size (in bytes). Higher values may help
  find string or regex patterns in very large pages, though doing so may imply
  more memory and CPU usage. The default value is 16384 and can be changed at
  build time. It is not recommended to change this value, but to use better
  checks whenever possible.

\subsubsection[tune.http.maxhdr]{tune.http.maxhdr <number>}
  Sets the maximum number of headers in a request. When a request comes with a
  number of headers greater than this value (including the first line), it is
  rejected with a "400 Bad Request" status code. Similarly, too large responses
  are blocked with "502 Bad Gateway". The default value is 101, which is enough
  for all usages, considering that the widely deployed Apache server uses the
  same limit. It can be useful to push this limit further to temporarily allow
  a buggy application to work by the time it gets fixed. Keep in mind that each
  new header consumes 32bits of memory for each session, so don't push this
  limit too high.

\subsubsection[tune.maxaccept]{tune.maxaccept <number>}
  Sets the maximum number of consecutive accepts that a process may perform on
  a single wake up. High values give higher priority to high connection rates,
  while lower values give higher priority to already established connections.
  This value is limited to 100 by default in single process mode. However, in
  multi-process mode (nbproc > 1), it defaults to 8 so that when one process
  wakes up, it does not take all incoming connections for itself and leaves a
  part of them to other processes. Setting this value to -1 completely disables
  the limitation. It should normally not be needed to tweak this value.

\subsubsection[tune.maxpollevents]{tune.maxpollevents <number>}
  Sets the maximum amount of events that can be processed at once in a call to
  the polling system. The default value is adapted to the operating system. It
  has been noticed that reducing it below 200 tends to slightly decrease
  latency at the expense of network bandwidth, and increasing it above 200
  tends to trade latency for slightly increased bandwidth.

\subsubsection[tune.maxrewrite]{tune.maxrewrite <number>}
  Sets the reserved buffer space to this size in bytes. The reserved space is
  used for header rewriting or appending. The first reads on sockets will never
  fill more than bufsize-maxrewrite. Historically it has defaulted to half of
  bufsize, though that does not make much sense since there are rarely large
  numbers of headers to add. Setting it too high prevents processing of large
  requests or responses. Setting it too low prevents addition of new headers
  to already large requests or to POST requests. It is generally wise to set it
  to about 1024. It is automatically readjusted to half of bufsize if it is
  larger than that. This means you don't have to worry about it when changing
  bufsize.

\subsubsection[tune.pipesize]{tune.pipesize <number>}
  Sets the kernel pipe buffer size to this size (in bytes). By default, pipes
  are the default size for the system. But sometimes when using TCP splicing,
  it can improve performance to increase pipe sizes, especially if it is
  suspected that pipes are not filled and that many calls to splice() are
  performed. This has an impact on the kernel's memory footprint, so this must
  not be changed if impacts are not understood.

\subsubsection[tune.rcvbuf.client]{tune.rcvbuf.client <number>}
\subsubsection[tune.rcvbuf.server]{tune.rcvbuf.server <number>}
  Forces the kernel socket receive buffer size on the client or the server side
  to the specified value in bytes. This value applies to all TCP/HTTP frontends
  and backends. It should normally never be set, and the default size (0) lets
  the kernel autotune this value depending on the amount of available memory.
  However it can sometimes help to set it to very low values (eg: 4096) in
  order to save kernel memory by preventing it from buffering too large amounts
  of received data. Lower values will significantly increase CPU usage though.

\subsubsection[tune.sndbuf.client]{tune.sndbuf.client <number>}
\subsubsection[tune.sndbuf.server]{tune.sndbuf.client <number>}
  Forces the kernel socket send buffer size on the client or the server side to
  the specified value in bytes. This value applies to all TCP/HTTP frontends
  and backends. It should normally never be set, and the default size (0) lets
  the kernel autotune this value depending on the amount of available memory.
  However it can sometimes help to set it to very low values (eg: 4096) in
  order to save kernel memory by preventing it from buffering too large amounts
  of received data. Lower values will significantly increase CPU usage though.
  Another use case is to prevent write timeouts with extremely slow clients due
  to the kernel waiting for a large part of the buffer to be read before
  notifying haproxy again.

\section{Debugging}

\subsubsection[debug]{debug}
  Enables debug mode which dumps to stdout all exchanges, and disables forking
  into background. It is the equivalent of the command-line argument "-d". It
  should never be used in a production configuration since it may prevent full
  system startup.

\subsubsection[quiet]{quiet}
  Do not display any message during startup. It is equivalent to the command-
  line argument "-q".


\section{Userlists}
It is possible to control access to frontend/backend/listen sections or to
http stats by allowing only authenticated and authorized users. To do this,
it is required to create at least one userlist and to define users.

\subsubsection[userlist]{userlist <listname>}
  Creates new userlist with name <listname>. Many independent userlists can be
  used to store authentication \& authorization data for independent customers.

\subsubsection[group]{group <groupname> [users <user>,<user>,(...)]}
  Adds group <groupname> to the current userlist. It is also possible to
  attach users to this group by using a comma separated list of names
  proceeded by "users" keyword.

\subsubsection[user]{user <username> [password | insecure-password <password>]
  [groups <group>, (...)]}
  Adds user <username> to the current userlist. Both secure (encrypted) and
  insecure (unencrypted) passwords can be used. Encrypted passwords are
  evaluated using the crypt(3) function so depending of the system's
  capabilities, different algorithms are supported. For example modern Glibc
  based Linux system supports MD5, SHA-256, SHA-512 and of course classic,
  DES-based method of crypting passwords.

  Example:
\begin{verbatim}
        userlist L1
          group G1 users tiger,scott
          group G2 users xdb,scott

          user tiger password $6$k6y3o.eP$JlKBx9za9667qe4(...)xHSwRv6J.C0/D7cV91
          user scott insecure-password elgato
          user xdb insecure-password hello

        userlist L2
          group G1
          group G2

          user tiger password $6$k6y3o.eP$JlKBx(...)xHSwRv6J.C0/D7cV91 groups G1
          user scott insecure-password elgato groups G1,G2
          user xdb insecure-password hello groups G2
\end{verbatim}

  Please note that both lists are functionally identical.


\section{Peers}

It is possible to synchronize server entries in stick tables between several
haproxy instances over TCP connections in a multi-master fashion. Each instance
pushes its local updates and insertions to remote peers. Server IDs are used to
identify servers remotely, so it is important that configurations look similar
or at least that the same IDs are forced on each server on all participants.
Interrupted exchanges are automatically detected and recovered from the last
known point. In addition, during a soft restart, the old process connects to
the new one using such a TCP connection to push all its entries before the new
process tries to connect to other peers. That ensures very fast replication
during a reload, it typically takes a fraction of a second even for large
tables.

\subsubsection[peers]{peers <peersect>}
  Creates a new peer list with name <peersect>. It is an independant section,
  which is referenced by one or more stick-tables.

\subsubsection[peer]{peer <peername> <ip>:<port>}
  Defines a peer inside a peers section.
  If <peername> is set to the local peer name (by default hostname, or forced
  using "-L" command line option), haproxy will listen for incoming remote peer
  connection on <ip>:<port>. Otherwise, <ip>:<port> defines where to connect to
  to join the remote peer, and <peername> is used at the protocol level to
  identify and validate the remote peer on the server side.

  During a soft restart, local peer <ip>:<port> is used by the old instance to
  connect the new one and initiate a complete replication (teaching process).

  It is strongly recommended to have the exact same peers declaration on all
  peers and to only rely on the "-L" command line argument to change the local
  peer name. This makes it easier to maintain coherent configuration files
  across all peers.

  Example:
\begin{verbatim}
    peers mypeers
        peer haproxy1 192.168.0.1:1024
        peer haproxy2 192.168.0.2:1024
        peer haproxy3 10.2.0.1:1024

    backend mybackend
        mode tcp
        balance roundrobin
        stick-table type ip size 20k peers mypeers
        stick on src

        server srv1 192.168.0.30:80
        server srv2 192.168.0.31:80
\end{verbatim}